Peer influences in air travel behavior: A social autoregressive analysis

### 2015-03-27 article.Rmd
Added code to calculate matrix weights.

---
title: "Peer influences in air travel behavior: A social autoregressive analysis"
author: Gregory S. Macfarlane, Thomas A. Wall, Bingling Zhang, and Kari E. Watkins
bibliography: bibliography/bibliography.bib
csl: bibliography/elsevier-harvard.csl 
output:
  html_document:
    fig_caption: yes
    number_sections: yes
---

@@

Candidate Weighting Schema
-----------------------------------------
The social network data we retrieved from Facebook allow us to construct four
social weights matrices, which we describe in this section. To aid this
presentation, we must first establish some basic notation following elementary
set theory. The set of friends of an individual  is defined as . The cardinality
of this set is the number of friends, so therefore the quantity  represents the
number of friends that individuals  and  have in common.

@@

**Mutual Friendship**
This schema defines a matrix where the cell value is the number of friends  and
have in common (as in Gilbert & Karahalios 2009; He et al. 2012),

**Network Overlap**
This schema defines a matrix where the cell value is the relative overlap of
each  pair’s social network (as in Adamic & Adar 2003),

**Direct Network Overlap**
This schema is similar to the Network Overlap, except that we give a bonus to
pairs that are themselves friends,

>>>>

---
title: "Peer influences in air travel behavior: A social autoregressive analysis"
author: Gregory S. Macfarlane, Thomas A. Wall, Bingling Zhang, and Kari E. Watkins
bibliography: bibliography/bibliography.bib
csl: bibliography/elsevier-harvard.csl 
output:
  html_document:
    fig_caption: yes
    number_sections: yes
---
```{r setup}
library(knitr) # nice output tables

# knit options
opts_chunk$set(
  cache = TRUE, cache.path = "./.cache/",
  echo = TRUE, 
  fig.keep = TRUE, fig.path = "./.figs/"
)

library(dplyr, warn.conflicts = FALSE) # super data wrangling
library(igraph) # graph analysis
library(networkD3) # interactive graph visualizer
```

@@

Candidate Weighting Schema
-----------------------------------------
The social network data we retrieved from Facebook allow us to construct four
social weights matrices, which we describe in this section. To aid this
presentation, we must first establish some basic notation following elementary
set theory. The set of friends of an individual  is defined as . The cardinality
of this set is the number of friends, so therefore the quantity  represents the
number of friends that individuals  and  have in common.
```{r connections_frame}
friend_connections <- tbl_df(expand.grid(id_i = user$id, id_j = user$id)) %>%
  group_by(id_i, id_j)
```

@@

$$ w_{ij} =
  \begin{cases}
   1 & \text{if } i \in \phi_j \\
   0 & \text{if } i \notin \phi_j
  \end{cases} $$

```{r direct_friends_graph}
check_friends <- function(id_i, id_j, user, friends){
  sum(user$fbid[user$id == id_i] %in% friends$fbid[friends$id == id_j])
}

friend_connections <- friend_connections %>%
  mutate(direct = check_friends(id_i, id_j, user, friends))
```
**Mutual Friendship**
This schema defines a matrix where the cell value is the number of friends  and
have in common (as in Gilbert & Karahalios 2009; He et al. 2012),
```{r mutual_friends}
count_mutual_friends <- function(id_i, id_j, friends){
  length(intersect(
  # sets of friend ids
    friends$fbid[friends$id == id_i],
    friends$fbid[friends$id == id_j]
  ))
}

friend_connections <- friend_connections %>%
  mutate(mutual = count_mutual_friends(id_i, id_j, friends))
```

**Network Overlap**
This schema defines a matrix where the cell value is the relative overlap of
each  pair’s social network (as in Adamic & Adar 2003),
```{r network_overlap}
calculate_newtork_overlap <- function(id_i, id_j, friends){
  mutual <- length(intersect( 
    # sets of friend ids
    friends$fbid[friends$id == id_i],
    friends$fbid[friends$id == id_j]
  ))

  total <- length(union( 
    friends$fbid[friends$id == id_i],
    friends$fbid[friends$id == id_j]
  ))

  mutual/total
}

friend_connections <- friend_connections %>%
  mutate(overlap = calculate_newtork_overlap(id_i, id_j, friends))
```

**Direct Network Overlap**
This schema is similar to the Network Overlap, except that we give a bonus to
pairs that are themselves friends,
```{r direct_network_overlap}
friend_connections <- friend_connections %>%
  mutate( direct_overlap = overlap + direct )
```
